---
title: "CMTH642 - Assignment 3"
author: "Mohammed Amir"
date: "December 16, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("gmodels")
#install.packages("class")
library(class)
library(gmodels)
library(corrplot)
```

Upload data
```{r}
wine <- read.csv(file="D:/Big Data/CMTH642 - DATA ANALYTICS ADVANCED METHODS/ASSIGNMENT 3/winequality-white.csv",head=TRUE,sep=";")
```


# 1. Check data characteristics. Is there missing data?
```{r}
#is.na(wine)
wine[!complete.cases(wine),]
str(wine)
```

# 2. What is the correlation between the attributes other than wine quality?
```{r}
## remove the wine quality attribute
newdf <- wine[-12]
head(newdf)

## Correlate wine attributes excluding quality
wine_correlation <- cor(newdf)
wine_correlation

## plot wine attribute correlation
corrplot(wine_correlation, method="circle")

```

# 3. Graph the distribution of wine quality.
```{r}
hist(wine$quality)

```

# 4. Reduce the levels of rating for quality to three levels as high, medium and low
```{r}
rating <- function(quality)
{
    if(quality <= 4)
    {
        print ("Low")
    }
        else
        {
            if(quality >= 5 && quality <= 7)
            {
              print ("Medium")  
            }
              else
                {
                    print ("High")
                }
          }
}
## add rating attribute to the wine data frame
wine$rating <- mapply(rating,wine$quality)
head(wine)


```

# 5. Normalize the data set.
```{r}
# remove quality attribute from wine data frame before normalization as it is not numeric attribute
new_wine_attribute <- wine[-12]

# Normaization function
normalize <- function(x) 
    {
      return ( 
                ( (x - min(x)) / (max(x) - min(x)))
              ) 
    }

dfNorm <- as.data.frame(lapply(new_wine_attribute[1:11], normalize))
head(dfNorm)

```

# 6 Divide the data to training and testing groups.
```{r}

dfNorm_train <- dfNorm[1:3265,]    ## two third of data set for training
dfNorm_test  <- dfNorm[3266:4898,] ## one third of data set for testing
nrow(dfNorm_train)
nrow(dfNorm_test)

```

# 7 Use the KNN algorithm to predict the quality of wine using its attributes.
```{r}


dfNorm_train_labels <- wine[1:3265, 12]    ## predict the quality in column 12
dfNorm_test_labels  <- wine[3266:4898,12] ## predict the quality in column 12

prediction <- knn(train = dfNorm_train, test = dfNorm_test,cl = dfNorm_train_labels, k=10)
```

# 8 Evaluate the model performance.
```{r}
#install.packages("gmodels")

CrossTable(x=dfNorm_test_labels, y = prediction, prop.chisq=FALSE)

```

