---
title: "CMTH642 Data Analytics Advanced Method - Assignment 3"
author: "Mohammed Amir"
date: "December 16, 2016"
output: word_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(class)
library(gmodels)
library(corrplot)
```
Upload data
```{r}
wine_df <- read.csv(file="D:/Big Data/CMTH642 - DATA ANALYTICS ADVANCED METHODS/ASSIGNMENT 3/winequality-white.csv",head=TRUE,sep=";")
```
# 1. Check data characteristics. Is there missing data?
```{r}
is.na(wine_df)
wine_df[!complete.cases(wine_df),]
str(wine_df)
summary(wine_df)
head(wine_df)
table(wine_df$quality)


```

# 2. What is the correlation between the attributes other than wine_df quality?
```{r}
## remove the wine_df quality attribute
new_wine_df <- wine_df[-12]
head(new_wine_df)

## Correlate wine_df attributes excluding quality
wine_df_correlation <- cor(new_wine_df)
wine_df_correlation

## plot wine_df attribute correlation
corrplot(wine_df_correlation, method="circle")

```

# 3. Graph the distribution of wine quality.
```{r}
hist(wine_df$quality)
```

# 4. Reduce the levels of rating for quality to three levels as high, medium and low
```{r}
rating <- function(quality)
{
    if(quality <= 4)
    {
        print ("Low")
    }
        else
        {
            if(quality >= 5 && quality <= 7)
            {
              print ("Medium")  
            }
              else
                {
                    print ("High")
                }
          }
}
## add rating attribute to the wine_df data frame
wine_df$rating <- mapply(rating,wine_df$quality)

# Total Count for each rating
 table(wine_df$rating)
```


# 5. Normalize the data set.
```{r}
# remove quality attribute from wine dataframe before normalization as it is not numeric attribute
new_wine_df_attribute <- wine_df[-12]

# Normaization function
normalize <- function(x) 
    {
      return ( 
                ( (x - min(x)) / (max(x) - min(x)))
              ) 
    }
winedf_Norm <- as.data.frame(lapply(new_wine_df_attribute[1:11], normalize))
head(winedf_Norm)
summary(winedf_Norm)
```

# 6 Divide the data to training and testing groups
```{r}
 ## Randomly divide the datasett to two sections 67% to 33%  
ind <- sample(2, nrow(winedf_Norm), replace=TRUE, prob=c(0.67, 0.33))
winedf_training <- winedf_Norm[ind==1,] ## assign 0.67 to training
winedf_test <- winedf_Norm[ind==2, ]     ## assign 0.33 to test 

nrow(winedf_training)  # number of rows in training dataset
nrow(winedf_test)      # number of rows in testing dataset
```

# 7 Use the KNN algorithm to predict the quality of wine_df using its attributes.
```{r}

winedf_training_Labels <- new_wine_df_attribute[ind==1, 12] ## Create a train labels
winedf_test_Labels     <- new_wine_df_attribute[ind==2, 12]  ## Create a test labels

#Apply a KNN function for k = 5
wine_prediction <- knn(train = winedf_training, test = winedf_test, cl = winedf_training_Labels, k=5)

```

# 8 Evaluate the model performance.
```{r}
CrossTable(x=winedf_test_Labels, y = wine_prediction, prop.chisq=FALSE)
```

