---
title: "CMTH642 - Assignment #2 "
author: "Mohammed Amir"
date: "November 26, 2016"
output: word_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read the csv files in the folder
```{r}
USDAClean <- read.csv(file="D:/Big Data/CMTH642 - DATA ANALYTICS ADVANCED METHODS/ASSIGNMENT 2/USDAClean_MohammedAmir.csv",head=TRUE,sep=",")
```

# 1. Create a visualization to illustrate the distribution of values for Total Fat, Protein and Carbohydrate. (12 p)
```{r}
TotalFat <- USDAClean$TotalFat
Protein <- USDAClean$Protein
Carbohydrate <- USDAClean$Carbohydrate

layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))
hist(TotalFat)
hist(Protein)
hist(Carbohydrate)
```

# 2. Create a visualization to illustrate the relationship between a food's Total Fat content and its calorie content. (12 p)
```{r}
# Based on the graphy below the relationship between Calories and Total Fat shows that, with lower total fat, the calories value concentration is high and as the total fat value increases the calories concentration starts to get less.

Calories <- USDAClean$Calories
plot (TotalFat, Calories,  col = c("green", "red"), xlab="TotalFat", ylab="Calories", main="Total Fat vs Calories realtionship graphy")
legend("topright", legend=c("TotalFat", "Calories"), col=c("green", "red"), lty=1:2, cex=1, box.lty=0)

```

# 3. Create a logistic regression model, using High Calories as the dependent variable, and Carbohydrate, Protein, Total Fat and Sodium as independent variables. (18 p)
```{r}
Sodium <- USDAClean$Sodium
logreg_model <- glm (HighCalories ~ Carbohydrate + Protein + TotalFat + Sodium, data=USDAClean, family = binomial, )
summary(logreg_model)

```
# 4. Which independent variable is not significant? (10 p)
```{r}
# As per the logistic regression value Sodium is not significant because it is p value of 0.347 is greater than 0.05 and has no star rating associated to it.
```


# 5. Which independent variable has the strongest positive predictive power in the model? (10 p)
```{r}
# Based on the highest coefficient number of TotalFat, it has the highest positive predictive value

```

# 6. Create a script for a HealthCheck function to detect unhealthy foods. (18 p) 
```{r}
HealthCheck <- function(sodium, sugar, fat)
{
    if(sodium == 0)
    {
        print ("1")
    }
        else
        {
            if(sugar == 0 && sodium == 1)
            {
              print ("1")  
            }
              else
                {
                  if(fat == 0 && sugar == 1 && sodium == 1)
                    {
                      print ("1")  
                    }
                      else
                        {
                          print ("0")
                        }
                
                }
          }
  }


```
# 7. Add a new column called HealthCheck to the USDAclean data frame using the output of the function. (10 p)
```{r, echo="FALSE"}

USDAClean$HealthCheck <- mapply(HealthCheck,USDAClean$HighSodium,USDAClean$HighSugar,USDAClean$TotalFat)

```
# 8. How many foods in the USDAcle an data frame fail the HealthCheck? (10 p)
```{r}
# 512 foods have failed the health check test were as 6101 foods have passed the test.

table(USDAClean$HealthCheck)
#head(USDAClean)
#write.csv(USDAClean, file = "D:/USDA_3.csv", row.names= TRUE)
#USDAClean$HealthCheck

```
